A tutorial on the model-dependent analysis of neutron and X-ray reflectometry

### 2020-08-07 refl_maths/paper.tex

% I know that it is common/precedent to start by looking at the Born approximation, but I wonder if, in paper format, it is just worth pointing out that multiple reflections invalidate Born and something iterative needs to be done? J.C.
%ARJN I've never looked at the BA first (apart from possibly looking at the FT briefly).

@@

The collected data (once normalised) is described in terms of reflected intensity, $R(q)$, which depends on the wavevector, $q$, and is measured as,
%
%ARJN, how about how to calculate the wavevector from a vector sum? i.e. something similar to fig 2, augmented with ki, kf.
\begin{equation}
    R(q) = \frac{\text{specular reflected intensity at $q$}}{\text{incident radiation intensity}},
    \label{equ:refl}
\end{equation}
%
where the denominator in Equation~\ref{equ:refl} is the total flux of X-rays or neutrons on the sample.
It is this $R(q)$ that we want to calculate from our structural model and then compare with our data.

@@

\section{Dynamical approach}
%ARJN abeles is not recursive
The breakdown of the kinematic approach leads to the application of the Abel\`{e}s \cite{abeles_sur_1948} or Parratt \cite{parratt_surface_1954}, recursive model for the reflection of light at a given number of stratified interfaces, also referred to as the dynamical approach.
This is an alternative method for calculating the reflected intensity from a model, however, now the model is a series of layers of different scattering length density.
The probing radiation can be either reflected or refracted at the interface between each of the layers.
%
\begin{figure}[t]
    \includegraphics[width=\linewidth]{dynamic}
    \caption{A schematic diagram showing the reflected ($r$) and transmitted ($t$) waves when an incident ($i$) wave enters an interface of thickness $d$.}
    \label{fig:refr}
\end{figure}
%
Figure~\ref{fig:refr} shows the refraction and reflections from a simple two layer system, where layer \num{0} is a vacuum above the sample.
It is clear, by analogy to Bragg's law, how the two waves, labelled $r$, could interfere constructively or destructively depending on the thickness of layer \num{1}, $d$.

>>>>

% I know that it is common/precedent to start by looking at the Born approximation, but I wonder if, in paper format, it is just worth pointing out that multiple reflections invalidate Born and something iterative needs to be done? J.C.
%ARJN I've never looked at the BA first (apart from possibly looking at the FT briefly).
%ARM I know that in Sivia (Elementary Scattering Theory) the BA/FT is introduced first, and I suppose that the context of this is that the previous chapter introduces scattering in the language of a FT. I am not religious to BA/FT coming first, but I think it is good from a historical perspective given that some paper use this appoarch.

@@

The collected data (once normalised) is described in terms of reflected intensity, $R(q)$, which depends on the wavevector, $q$, and is measured as,
%
%ARJN, how about how to calculate the wavevector from a vector sum? i.e. something similar to fig 2, augmented with ki, kf.
%ARM Do you mean something like the $|k_i|\sin{\theta} = |k_f|\sin{\theta} = \frac{2\pi}\sin{\theta} = \frac{|\mathbf{q}|}{2} = \frac{q}{2}$ as a why of explaining what the $q$-vector is? I agree that it is necessary to explain what $q$ is but how about we just give the $q = \frac{4\pi\sin{\theta}}{\lambda} and explain that a full derivation can be found elsewhere \cite{sivia_elementary_2011}
\begin{equation}
    R(q) = \frac{\text{specular reflected intensity at $q$}}{\text{incident radiation intensity}},
    \label{equ:refl}
\end{equation}
%
where the denominator in Equation~\ref{equ:refl} is the total flux of X-rays or neutrons on the sample.
It is this $R(q)$ that we want to calculate from our structural model and then compare with our data.

@@

\section{Dynamical approach}
The breakdown of the kinematic approach leads to the application of the Abel\`{e}s matrix formalism \cite{abeles_sur_1948} or Parratt \cite{parratt_surface_1954} recursive model for the reflection of light at a given number of stratified interfaces, also referred to as the dynamical approach.
This is an alternative method for calculating the reflected intensity from a model, however, now the model is a series of layers of different scattering length density.
The probing radiation can be either reflected or refracted at the interface between each of the layers.
%
\begin{figure}[t]
    \includegraphics[width=\linewidth]{dynamic}
    \caption{A schematic diagram showing the reflected ($r$) and transmitted ($t$) waves when an incident ($i$) wave enters an interface of thickness $d$.}
    \label{fig:refr}
\end{figure}
%
Figure~\ref{fig:refr} shows the refraction and reflections from a simple two layer system, where layer \num{0} is a vacuum above the sample.
It is clear, by analogy to Bragg's law, how the two waves, labelled $r$, could interfere constructively or destructively depending on the thickness of layer \num{1}, $d$.
